---
title: "Case Study 2"
author: "Marc Naples"
date: "December 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Case Study 2 - Doing Data Science, Fall 2016

## QUestion 1 - Create Matrix and print it

### SAS Version

### R Version

```{r}
X = matrix( 
  c(4, 1, 2, 5, 0, 1, 1, 3, 8, 2, 5, 2), 
  nrow=3, 
  ncol=4)
X
```

### Python Version 
See notebook located on this GitHub project.

## Question 2 - Week 11 Lecture Assignment

Download the Data

```{r}
library(tseries)

TCPdata <- get.hist.quote('TCP',quote="Close")
```

Calculate Log Returns

```{r}
TCPret <- log(lag(TCPdata)) - log(TCPdata)
TCPvol <- sd(TCPret) * sqrt(250) * 100
```

Calculate Volatility Measure

```{r}
getVol <- function(d, logrets)
{
  var = 0
  lam = 0
  varlist <- c()
  for (r in logrets) {
    lam = lam*(1 - 1/d) + 1
    var = (1 - 1/lam)*var + (1/lam)*r^2
    varlist <- c(varlist, var)
  }
  sqrt(varlist)
}
```

Calculate volatility over entire length of series for the different decay factors

```{r}
volest <- getVol(10,TCPret)
volest2 <- getVol(30,TCPret)
volest3 <- getVol(100,TCPret)
```


Plot the Results

```{r}
plot(volest,type="l")
lines(volest2,type="l",col="red")
lines(volest3, type = "l", col="blue")
```

## Question 3 - Orange Data Problems

### Calculate the mean and the median of the trunk circumference for different size of the trees (Tree)

```{r}
# Import data
library(datasets)
data(Orange)

# Calculations
tapply(Orange$circumference, Orange$Tree, mean)
tapply(Orange$circumference, Orange$Tree, median)
```

### Make a scatterplot of the trunk circumferences against the age of the tree. Use different plotting symbols for different size of trees

```{r}
# Reorder Data
Orange$order <- as.factor(rep(c(3,1,5,2,4), each=7))

# Plot
library(ggplot2)
ggplot(Orange, aes(x=age, y=circumference)) + 
  geom_point(size=2, shape=23, aes(fill=Orange$Tree)) + xlab("Age") + ylab("Circumference") + labs("Tree Type)") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), legend.title=element_blank()) + ggtitle("Circumference Vs. Age")
```


### Display the trunk circumference on a comparative boxplot against tree (ensure boxplots ordered such that they increase with maximum diameter)

```{r}
boxplot(circumference ~ Tree, data = Orange)
title("Boxplot of Trunk Circumference, by Tree Diameter Type")
```

## Question 4 - Problems with Temp dataset

### Find the idfference between the maximum and minimum monthly average temps for each country and report/visualize top 20 countires with the maximum differences for the period since 1900.

```{r}
# Import temp data
setwd('C:/R/CaseStudy2')
temp <- read.csv('temp.csv', header=TRUE)
library(plyr)

## clean data set by remove NAs, rename 'date' to 'month'
cleantemp <- na.omit(temp)
names(cleantemp)[1] <- "month"

## Change month column to date type object with step that only works on dates entered after 1900, remove NAs (thus removing pre-1900 rows)
cleantemp$month <- as.Date(as.character(cleantemp$month), "%m/%d/%Y")
cleantemp <- na.omit(cleantemp)

## calculate mins and maxes for each Country, create new data set with values
maxes <- aggregate(cleantemp$Monthly.AverageTemp, by=list(cleantemp$Country), FUN=max)
mins <- aggregate(cleantemp$Monthly.AverageTemp, by=list(cleantemp$Country), FUN=min)
range <- cbind(maxes, mins)
names(range)[2] <- "max"
names(range)[4] <- "min"

## add column for difference between min and max, sort descending by range, keep only top 20
range$diff <- range$max - range$min
range.sorted <- range[order(-range$diff),]
range.sorted <- range.sorted[1:20, ]

## Plot
barplot(range.sorted$diff, ylab="Temp Range", names.arg=range.sorted$Group.1, las=2)
```

### Select a subset of US tempatures since 1990

```{r}
library(dplyr)
UStemp <- subset(cleantemp, Country=="United States" & month >= as.Date("1990-01-01"), select=month:Monthly.AverageTemp)
```

#### Create a new column to display the monthly average land temps in Farhrenheit

```{r}
UStemp$ftemp <- with(UStemp, Monthly.AverageTemp * (9/5) + 32)
```

#### Calculate average land temperature by year and plot it

```{r}
# Create new column identifying year only of date
UStemp$year <- format(as.Date(UStemp$month, format="%Y-%m-%d"),"%Y")

# Calculate annual mean temps, store to object and convert to dataframe
annualtemps <- tapply(UStemp$ftemp, UStemp$year, mean)
annualdf <- as.data.frame.table(annualtemps)

# Clean data frame
annualdf$Var1 <- as.numeric(annualdf$Var1)
annualdf$Var1 <- with(annualdf, Var1 + 1989)
names(annualdf)[1] <- "Year"
names(annualdf)[2] <- "Temp"

# Plot
plot(annualdf, type="l", xlab=("Year"), ylab=("Average Annual Temperature (Fahrentheit)"))
```

#### Calculate the one year difference of average land temperature and provide maximum different between years

```{r}
# Calculate year to year differences from dataframe of annual temps created above
diff(annualdf$Temp)

# Store calculations, add as column of data to dataframe
tempchanges <- diff(annualdf$Temp)
annualdf$diff <- c(NA, tempchanges)
annualdf
```
Examination shows the greatest one year difference to occur in the final row of the dataframe, with temperatures rising from 2012 to 2013.

### Use CityTemp data, find the difference between maximum and minimum temps for each major city and report/visualize top 20 cities since 1900.

```{r}
# Import CityTemp
setwd('C:/R/CaseStudy2')
citytemp <- read.csv('CityTemp.csv', header=TRUE)

## clean data set by remove NAs, rename 'date' to 'month'
cleancity <- na.omit(citytemp)
names(cleancity)[1] <- "month"

## Change month column to date type object with step that only works on dates entered after 1900, remove NAs (thus removing pre-1900 rows)
cleancity$month <- as.Date(as.character(cleancity$month), "%m/%d/%Y")
cleancity <- na.omit(cleancity)

## calculate mins and maxes for each city, create new data set with values
citymaxes <- aggregate(cleancity$Monthly.AverageTemp, by=list(cleancity$City), FUN=max)
citymins <- aggregate(cleancity$Monthly.AverageTemp, by=list(cleancity$City), FUN=min)
cityrange <- cbind(citymaxes, citymins)
names(cityrange)[2] <- "max"
names(cityrange)[4] <- "min"

## add column for difference between min and max, sort descending by range, keep only top 20 results
cityrange$diff <- cityrange$max - cityrange$min
cityrange.sorted <- cityrange[order(-cityrange$diff),]
cityrange.sorted <- cityrange.sorted[1:20, ]

## Plot
barplot(cityrange.sorted$diff, ylab="Temp Range", names.arg=cityrange.sorted$Group.1, las=2)
```

#### Compare and comment on City versus Country temp differences



